Parsing Postfix Log Files

### 2009-04-22 doc/logparser-results.tex
Corrections after proofreading chapters 6 & 7 and the appendices.

This chapter evaluates \parsername{} on two criteria: efficiency, and
coverage of  Postfix log files.  \parsername{} is intended to be used in a
production environment, and must be capable of parsing log files generated
by a high volume mail server in a reasonable time period.  Accurate but
slow parsing is preferable to sloppy but quick parsing, so the goal of
better parser efficiency must be balanced against the requirement for
precise and correct processing of log files.

The performance evaluation begins by describing the characteristics of the
mail server that produced the log files used evaluate the parser's
performance and coverage; the computer that the tests were run on is also
described.  The characteristics of the \numberOFlogFILES{} log files are
described next, with an explanation of why \parsername{} has better
performance with the larger log files in the group.  The effect of using
different rule orderings is explored, and the use of optimal rule ordering
is compared to use of an oracle that allows the parser to use only one rule
when recognising each log line.  When \parsername{} is used by other mail
administrators, they will need to extend the ruleset to parse their own log
lines, so the next section addresses the question of how parser performance
is affected by an increase in the number of rules in the ruleset.  The
penultimate topic to be covered is the simple optimisation of caching the
results of compiling each regex, and the huge effect it has on parser
efficiency.  The performance evaluation concludes by examining where
parsing time is spent: recognition of log lines or their subsequent
processing?

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  This topic consists of two sections: what proportion of log
lines are correctly recognised by the ruleset, and what proportion of mail
delivery attempts are correctly understood and reconstructed by the
actions?  The former is a requirement for the latter to be achieved, and
the latter is important because the data provided by \parsername{} must be
both complete and correct for it to be of use to others.

@@

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The server that generated the log files used in testing
this parser accepts approximately 10,000 mails for 700 users each weekday;
\graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, far more mails are received on
weekdays than at weekends.  Note that these figures only count mails
received by \acronym{SMTP}, and the mail loops noticeable in later graphs
do not affect the figures.  Median log file size is
\input{build/include-median-file-size}, containing
\input{build/include-median-number-of-log-lines} log lines; large scale
mail servers would have much larger log files.  The mail server in question
is a production mail server handling mail for a university department; the
benefit of using this mail server is that its log files exhibit the
idiosyncrasies and peculiarities a mail server in the wild must deal with,
but the downside is that significantly altering the configuration to
accommodate this project is not an option.

@@

\begin{table}[thbp]
    \caption{Details of the computer used to generate statistics}
    \empty{}\label{Details of the computer used to generate statistics}
    \centering{}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component in use                                   \\
        \tablemiddleline{}%
        CPU         & One dual core 2.40GHz Intel\textregistered{}
                        Core\texttrademark{}~2 CPU,                     \\
                    & with 32KB L1 cache and 4MB L2 cache.              \\
        RAM         & 2GB 667 MHz DDR RAM\@.                            \\
        Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling \input{build/include-total-size-of-93-log-files},
\numberOFlogLINEShuman{} log lines) were each parsed 10 times, and the mean
parsing time.  The computer used for test runs was a Dell Optiplex 745,
shown in \tableref{Details of the computer used to generate statistics}; it
was dedicated to the task of gathering statistics from test runs, and did
not run any other programs while test runs were in progress.  Saving
results to the database was disabled for the test runs, because that
dominates the run time of the parser, and the tests are aimed at measuring
the speed of \parsername{} rather than the speed of the database and the
disks it is stored on.  Parsing all \numberOFlogFILES{} log files in one
run without saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing makes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.

@@

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  \Graphref{parsing time vs log file size vs number of log
lines graph} shows the parsing time in seconds, log file size in MB, and
number of log lines in tens of thousands, for each of the
\numberOFlogFILES{} log files.  \Graphref{parsing time vs log file size vs
number of log lines factor} plots both the ratio of log file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is more efficient in both cases); \tableref{parsing time vs log file size
vs number of log lines factor table} shows the same ratios for different
groups of log files.  The ratios are quite tightly banded, except they
increase (i.e.\ improve) for log files 22 \& 62--68, despite their larger
than usual size.  Both groups of log files are much larger than usual
because of mail loops caused by users who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loops have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the mean number of rules tried per
log line reduces, as does the mean parsing time per log line.
\Tableref{Number of rules per Postfix component} shows the number of rules
for each Postfix component; \graphref{Mean number of rules used per log
line} shows the drop in the mean number of rules tried per log line for log
files containing a mail loop, and \graphref{Mean number of rules used per
log line for each Postfix component} shows the mean number of rules tried
per log line for each Postfix component.

@@

\showgraph{build/graph-average-number-of-rules-tried-per-program}{Mean
number of rules used per log line for each Postfix component}{Mean number
of rules used per log line for each Postfix component}

@@

\parsernames{} ruleset has \numberOFrules{} rules: the top 10\% recognise
\input{build/include-top-ten-hits}\% of the log lines in the
\numberOFlogFILES{} log files, with the remaining log lines split across
the other 90\% of the rules, as shown in \graphref{rule hits graph}.
Assuming that the distribution of log lines is reasonably consistent over
time, \parsernames{} efficiency should benefit from using rules that
recognise log lines more frequently before it uses rules that recognise log
lines less frequently.  To test this hypothesis, three full test runs were
performed with different rule orderings:

@@

    \item [Optimal]  The most optimal order, according to the hypothesis:
        rules that recognise log lines most often will be tried first.

@@

    \item [Reverse] Hypothetically the worst order: rules that recognise
        log lines most frequently will be tried last.

@@

\Graphref{Parsing time of oracle and optimal ordering relative to shuffled
ordering graph} shows how the oracle and optimal ordering perform relative
to shuffled ordering; \tableref{Parsing time of oracle and optimal ordering
relative to shuffled ordering table} shows mean and standard deviation.  As
expected, the oracle is more efficient than optimal ordering, but not by
much.  \Graphref{Percentage increase in parsing time when using optimal
ordering instead of oracle graph} shows the percentage increase in parsing
time when using optimal ordering instead of the oracle, with mean and
standard deviation in \tableref{Percentage increase in parsing time when
using optimal ordering instead of oracle table}.

Once again, the difference between the oracle and optimal ordering is at
its lowest when parsing log files resulting from a mail loop (log files 22
\& 62--68), because the mean number of rules tried per log line is lower
(see \graphref{Mean number of rules used per log line}).  The performance
of optimal ordering relative to the oracle is much worse when parsing the
first log file than for the remainder of the log files, because for the
first log file the hits field of every rule is zero, so optimal ordering
does not provide any benefit for that log file; the oracle, in contrast, is
flawless for every log file.

@@

Clearly the increased number of rules causes a noticeable performance
decrease with reverse ordering, and a lesser decrease with shuffled
ordering, whereas optimal ordering shows scant change.  Log files 22 \&
62--68 show much smaller increases in parsing time than other log files do,
because most of the log lines in those log files are produced by Postfix
components with few rules, so removing unnecessary rules has little effect
on the total number of rules used; \tableref{Number of rules per Postfix
component in the maximum and minimum rulesets} shows the number of rules
per Postfix component for each ruleset.  Once again, for the first log
file, optimal and reverse orderings have identical performance, because the
hits field of every rule starts at zero.

The optimal ordering has a mean increase of just
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that both the architecture and \parsername{} scale
extremely well as the ruleset increases in size, and that optimally
ordering the rules is an important optimisation contributing to this
scalability.

@@

\begin{table}[thbp]
    \caption{Number of rules per Postfix component in the maximum and
    minimum rulesets}
    \empty{}\label{Number of rules per Postfix component in the maximum and
    minimum rulesets}
    \centering{}
    \begin{tabular}{lrr}
        \tabletopline{}%
        Postfix component & Maximum ruleset & Minimum ruleset \\
        \tablemiddleline{}%
        \input{build/include-number-of-rules-per-program-minimum-ruleset}
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and optimised
to improve the speed of matching.  This compilation and optimisation takes
CPU time: for most regexes it takes far more CPU time than the actual
matching.  If the interpreter is certain that a regex will not change while
the program is running, it will automatically cache the results of
compiling and optimising the regex for later use.  The results of compiling
a dynamically generated regex can be cached and used in preference to the
original regex, but it is the responsibility of the programmer to do so;
\parsername{} does this with every rule's regex when the rules are loaded
from the database.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the effect that not caching compiled regexes has on parser
performance, with mean and standard deviation in \tableref{Increase in
parsing time when not caching compiled regexes table}.  For typical log
files, the mean increase in parsing time when not caching compiled regexes
is \input{build/include-cached-regexes-vs-discarded-regexes-mean}.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement.  As seen
previously, log files 22 \& 62--68 do not suffer such a large increase in
parsing time when the optimisation is disabled; this is because, on
average, fewer regexes are compiled per log line for those log files.  The
increase in parsing time when parsing the first log file is much greater
than for the other log files; again, this is because every rule's hits
field starts at zero, so optimal ordering is less efficient than usual, the
mean number of rules tried for each log line will be higher than usual, and
more regexes will need to be compiled when recognising each log line.

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still dominates the execution time of the
        parser.  \Graphref{Percentage of parsing time spent recognising log
        lines graph} shows the percentage of parsing time spent recognising
        log lines for each of the \numberOFlogFILES{} log files, with mean
        and standard deviation shown in \tableref{Percentage of parsing
        time spent recognising log lines table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  Optimal ordering
        reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; processing recognised log lines
        occupies
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  If processing
        of recognised log lines was optimised to 1\% of its original
        parsing time, it would be only slightly more effective than optimal
        ordering, but it would be vastly harder to implement; optimising
        actions would not reduce parsing time enough to justify the amount
        of effort required.

@@

Individual actions or the framework could and have been optimised, but
plenty of existing literature is available on the topic of optimising
programs, so the subject will not be dealt with here.

@@

The discussion of \parsernames{} coverage of Postfix log files is separated
into two parts: log lines correctly recognised, and mail delivery attempts
correctly understood --- the former is a requirement for the latter to be
achieved.  Correctly understanding and reconstructing every mail delivery
attempt is important so that the information in the database is correct and
complete.  Improving the proportion of log lines correctly recognised is
the less difficult of the two, because it just requires new rules to be
written or existing rules to be extended.  Improving the proportion of
correctly understood and reconstructed mail delivery attempts is more
difficult and intrusive, because it requires adding or changing actions,
and it can be much harder to realise that a deficiency exists and needs to
be addressed.

@@

    \item Skip the log line if the ruleset does not contain any rules for
        the Postfix component that produced the log line.

    \item Try each rule until a recognising rule is found; if the log line
        is not recognised, issue a warning and move on to the next log
        line.

@@

Each Postfix component whose log lines are of interest must have at least
one rule, or its log lines will be silently skipped; in the extreme case of
an empty ruleset the parser would skip every log line.  \parsername{} skips
log lines logged by programs that do not have any rules because there may
be any number of log lines from other programs intermingled in the log
file, and some Postfix components that do not produce any log lines of
interest.  \parsername{} does not parse log lines from non-Postfix
programs, e.g.\ Amavisd-new or SpamAssassin; it could easily be extended to
do so, if a method could be developed to correctly associate such log lines
with existing state table entries.  To correctly recognise all log lines,
there must be a rule to recognise each log line variant produced by each
Postfix component; if a log line is not recognised the parser will issue a
warning, to inform the user that they need to extend their ruleset.  Each
rule's regex should be as specific and precise as possible, to ensure
accurate parsing: a rule with a regex that matches zero or more of any
character will recognise every log line, but not in a meaningful way; most
log lines contain fixed strings, so this is not a problem in practice.

Full coverage of log lines can be achieved without undue effort, yet once
achieved it is hard to maintain.  Maintaining full coverage is hard because
the set of log line variants changes over time, e.g.\ administrators add
restrictions with custom messages, \acronym{DNSBL} messages change, or
major releases of Postfix change log lines (usually by adding more
information).  Warnings are issued for any log lines that are not
recognised; no warnings are issued for unrecognised log lines while parsing
the \numberOFlogFILES{} log files, so it can be safely concluded that zero
false negatives arise.  False positives are harder to quantify, short of
examining each of the 60,721,709 log lines and checking that the correct
rule recognised it.  However, a random sample of 6039 log lines was parsed,
and the results manually verified by inspection to ensure that the correct
rule recognised each log line.  The sample was generated by running the
command: \texttt{perl -n -e \singlequote{}print if (rand 1 <
0.0001)\singlequote{} LOG\_FILES} to randomly extract roughly one log line
in every 10,000 (it actually extracted 0.00994\% instead of 0.01\%).  Each
log line was examined and the correct rule identified from the
\numberOFrules{} rules in the database; the correct rule was then compared
to the rule that recognised the log line when parsing.  The sample results
contained zero false positives, and this check has been automated to ensure
continued accuracy.  Based on these results, and how precise each rule's
regex is, the author is confident that zero false positives occur when
parsing the \numberOFlogFILES{} log files.  On initial appearances,
exercising only 36 rules from a total of \numberOFrules{} when parsing 6039
log lines seems low, but after examining \graphref{rule hits graph} it
becomes apparent that such a low number of rules is to be expected.  The
reader should also bear in mind that even when parsing all
\numberOFlogFILES{} log files, only \numberOFrulesMINIMUM{} rules are used,
because many of the rules recognise log lines that only appear in other log
files.

@@

The proportion of mail delivery attempts that are correctly understood and
reconstructed is much more difficult to determine accurately than the
proportion of log lines that are correctly recognised.  The parser can dump
its state tables in a human readable form; examining those tables with
reference to the log files is the best way to detect mails that were not
handled properly (many of the complications discussed in
\sectionref{complications} were detected in this way).  \parsername{}
issues warnings when it detects any errors or discrepancies, e.g.\ when a
queueid is reused but the previous mail remains in the state tables, when a
queueid or \acronym{pid} is not found in the state tables, or when an entry
in the state tables does not include sufficient data to satisfy the
database schema.  The parser should produce few or no warnings during
parsing, and when finished parsing the state tables should only contain
entries for mails that have log lines in subsequent log files.  There will
often be warnings about a missing queueid or \acronym{pid} when parsing the
first few thousand log lines, because the earlier log lines for those
connections or mails are in a previous log file; loading saved state tables
from the previous log file will solve this problem.

@@

The state tables will contain entries for mails not yet delivered when the
parser finishes execution.  Ideally, that is all they will contain, though
they may also contain mails whose initial log lines are not contained in
the log files.  Any other entries in the state tables are evidence of
either a failure in parsing, or an aberration in the log files.  After
parsing the \numberOFlogFILES{} log files, the state tables contain 18
entries:

@@

    \item 1 connection that started only seconds before the log files
        ended and had not yet been fully transferred from client to server.

@@

    \item \parsername{} performs many checks to detect known problems
        occurring, e.g.\ if a mail is missing required data, and no
        warnings are produced during the test runs other than the five
        described above.

@@

    \item When dealing with the complications described in
        \sectionref{complications}, the solutions are as specific and
        restrictive as possible, with the goal of minimising the number of
        false positives.  In addition, the solution to the complication
        described in \sectionref{out of order log lines} imposes conditions
        that each reconstructed mail must comply with to be acceptable.

>>>>

This chapter evaluates \parsername{} on two criteria: efficiency, and
coverage of  Postfix log files.  Efficiency is important because
\parsername{} is intended to be used in a production environment, and must
be capable of parsing log files generated by a high volume mail server in a
reasonable period of time.  Full coverage of Postfix log files is important
because the data gathered by the parser must be accurate and complete for
it to be useful and reliable.  The two goals are mutually incompatible:
every extra measure implemented to improve accuracy slows down parsing.
Some of the complications described in \sectionref{complications} occur
fewer than ten times when parsing the \numberOFlogFILES{} log files used to
evaluate \parsername{} in this chapter, yet their solutions slow down
parsing of every log file.  However, accurate but slow parsing is
preferable to sloppy but quick parsing.

The performance evaluation begins by describing the characteristics of the
mail server that produced the log files used to evaluate the parser's
performance and coverage; the computer that the tests were run on is also
described.  The characteristics of the \numberOFlogFILES{} log files are
described next, with an explanation of why \parsername{} has better
performance when parsing the larger log files in the group.  The effect of
using different rule orderings is explored, and the use of optimal rule
ordering is compared to use of an oracle that allows the parser to use only
one rule when recognising each log line.  When \parsername{} is used by
other mail administrators, they will need to extend the ruleset to parse
their own log lines, so the next section addresses the question of how
parser performance is affected by an increase in the number of rules in the
ruleset.  The penultimate topic covered is the simple optimisation of
caching the results of compiling each rule's regex, and the huge effect it
has on parser efficiency.  The performance evaluation concludes by
examining where parsing time is spent: recognition of log lines or their
subsequent processing?

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  This subject contains two sections: what proportion of log
lines are correctly recognised by the ruleset, and what proportion of mail
delivery attempts are correctly understood and reconstructed by the
actions?  The former is a requirement for the latter to be achieved, and
the latter is important because the data provided by \parsername{} must be
both complete and correct for it to be of use to others.

@@

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files is a production mail server handling mail for a university
department; the benefit of using this mail server is that its log files
exhibit the idiosyncrasies and peculiarities that a mail server in the wild
must deal with, but the downside is that significantly altering the
configuration to accommodate this project is not an option.
\Graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, it accepts far more mails on
weekdays than at weekends.  Note that these figures only count mails
received by \acronym{SMTP}, and the mail loops noticeable in later graphs
do not affect the figures.

@@

\begin{table}[thbp]
    \caption{Details of the computer used to generate statistics}
    \empty{}\label{Details of the computer used to generate statistics}
    \centering{}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component in use                                   \\
        \tablemiddleline{}%
        CPU         & One dual core 2.40GHz Intel\textregistered{}
                        Core$^{\textsc{TM}}$2 CPU,                      \\
                    & with 32KB L1 cache and 4MB L2 cache.              \\
        RAM         & 2GB 667 MHz DDR RAM\@.                            \\
        Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling \input{build/include-total-size-of-93-log-files},
\numberOFlogLINEShuman{} log lines) were each parsed 10 times, and the mean
parsing time calculated.  The computer used for test runs was a Dell
Optiplex 745, shown in \tableref{Details of the computer used to generate
statistics}; it was dedicated to the task of gathering statistics from test
runs, and did not run any other programs while test runs were in progress.
Saving results to the database was disabled for the test runs, because that
dominates the run time of the parser, and the tests are aimed at measuring
the speed of \parsername{} rather than the speed of the database and the
disks it is stored on.  Parsing all \numberOFlogFILES{} log files in one
run without saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.

@@

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  Median log file size is
\input{build/include-median-file-size}, containing
\input{build/include-median-number-of-log-lines} log lines, with a median
of \input{build/include-median-mails-accepted-over-smtp} mails accepted
each weekday; large scale mail servers would generate much larger log
files.  \Graphref{parsing time vs log file size vs number of log lines
graph} shows the parsing time in seconds, log file size in MB, and number
of log lines in tens of thousands, for each of the \numberOFlogFILES{} log
files.  \Graphref{parsing time vs log file size vs number of log lines
factor} plots both the ratio of log file size vs parsing time, and the
ratio of number of log lines vs parsing time (higher is more efficient in
both cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows the same ratios for different groups of log files.  The
ratios are quite tightly banded, except they increase (i.e.\ improve) for
log files 22 \& 62--68, despite their larger than usual size.  Both groups
of log files are much larger than usual because of mail loops caused by
users who set up mail forwarding incorrectly, resulting in a very different
distribution of log lines: normally most log lines are generated by mail
delivery attempts from other hosts, but when the mail loops occurred most
of the log lines resulted from failed delivery of mail generated on the
server itself.  The Postfix components that generated most of the log lines
during the mail loops have fewer associated rules than the Postfix
components whose log lines normally make up the bulk of each log file, so
the mean number of rules tried per log line is lower for these log files,
as is the mean parsing time per log line.  \Tableref{Number of rules per
Postfix component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules tried per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules tried per log line for each
Postfix component for all log files.

@@

\showgraph{build/graph-average-number-of-rules-tried-per-program}{Mean
number of rules used per log line for each Postfix component over all log
files}{Mean number of rules used per log line for each Postfix component}

@@

\parsernames{} ruleset has \numberOFrules{} rules: the top 10\% recognise
\input{build/include-top-ten-hits}\% of the log lines in the
\numberOFlogFILES{} log files, with the remaining log lines split across
the other 90\% of the rules, as shown in \graphref{rule hits graph}.
Assuming that the distribution of log lines is reasonably consistent across
log files, \parsernames{} efficiency should benefit from using rules that
recognise log lines more frequently before it uses rules that recognise log
lines less frequently.  To test this hypothesis, three full test runs were
performed with different rule orderings:

@@

    \item [Optimal]  The most optimal order, according to the hypothesis:
        rules that recognise log lines most often will be used first.

@@

    \item [Reverse] Hypothetically the worst order: rules that recognise
        log lines most frequently will be used last.

@@

\Graphref{Parsing time of oracle and optimal ordering relative to shuffled
ordering graph} shows how the oracle and optimal ordering perform relative
to shuffled ordering, with mean and standard deviation in \tableref{Parsing
time of oracle and optimal ordering relative to shuffled ordering table}.
As expected, the oracle is more efficient than optimal ordering, but not by
much.  \Graphref{Percentage increase in parsing time when using optimal
ordering instead of oracle graph} shows the percentage increase in parsing
time when using optimal ordering instead of the oracle, with mean and
standard deviation in \tableref{Percentage increase in parsing time when
using optimal ordering instead of oracle table}.

Once again, the difference between the oracle and optimal ordering is at
its lowest when parsing log files resulting from a mail loop (log files 22
\& 62--68), because the mean number of rules used per log line is lower
(see \graphref{Mean number of rules used per log line}).  The performance
of optimal ordering relative to the oracle is much worse when parsing the
first log file than for the remainder of the log files, because the hits
field of every rule starts at zero, so optimal ordering does not provide
any benefit for the first log file; the oracle, in contrast, is flawless
for every log file.

@@

Clearly the increased number of rules causes a noticeable performance
decrease with reverse ordering, and a lesser decrease with shuffled
ordering, whereas optimal ordering shows scant change.  Log files 22 \&
62--68 show much smaller increases in parsing time than other log files do,
because most of the log lines in those log files are produced by Postfix
components with few rules, so removing unnecessary rules has little effect
on the total number of rules used; \tableref{Number of rules per Postfix
component in the maximum and minimum rulesets} shows the number of rules
per Postfix component for each ruleset.  Once again, \graphref{Percentage
parsing time increase when using the maximum ruleset instead of the minimum
ruleset} shows that optimal and reverse orderings have identical
performance for the first log file, because the hits field of every rule
starts at zero.

The optimal ordering shows a mean increase of just
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that both the architecture and \parsername{} scale
extremely well as the ruleset increases in size, and that optimally
ordering the rules is an important optimisation contributing to this
scalability.

@@

\begin{table}[thbp]
    \caption{Number of rules per Postfix component in the maximum and
    minimum rulesets}
    \empty{}\label{Number of rules per Postfix component in the maximum and
    minimum rulesets}
    \centering{}
    \begin{tabular}{lrrr}
        \tabletopline{}%
        Postfix component & Maximum ruleset & Minimum ruleset & Difference \\
        \tablemiddleline{}%
        \input{build/include-number-of-rules-per-program-minimum-ruleset}
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and optimised
to improve the speed of matching.  This compilation and optimisation takes
CPU time: for most regexes it takes far more CPU time than the actual
matching does.  If the interpreter is certain that a regex will not change
while the program is running, it will automatically cache the results of
compiling and optimising the regex for later use.  The results of compiling
a dynamically generated regex can be cached and used in preference to the
original regex, but it is the responsibility of the programmer to do so;
\parsername{} does this with every rule's regex when the rules are loaded
from the database.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the effect that not caching compiled regexes has on parser
performance, with mean and standard deviation in \tableref{Increase in
parsing time when not caching compiled regexes table}.  For typical log
files, the mean increase in parsing time when not caching compiled regexes
is \input{build/include-cached-regexes-vs-discarded-regexes-mean}.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement.  As seen
previously, log files 22 \& 62--68 do not suffer such a large increase in
parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used so fewer regexes are compiled per log line
for those log files.  The increase in parsing time when parsing the first
log file is much greater than for the other log files; again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  Optimal ordering
        reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  If processing
        of recognised log lines was optimised to 1\% of its original
        parsing time, it would be only slightly more effective than optimal
        ordering, but it would be vastly harder to implement; optimising
        actions would not reduce parsing time enough to justify the amount
        of effort required.

@@

Individual actions or the framework could and have been optimised, but
plenty of existing literature is available on the topic of optimising
programs, so the subject is not dealt with here.

@@

The discussion of \parsernames{} coverage of Postfix log files is separated
into two parts: log lines correctly recognised, and mail delivery attempts
correctly understood --- the former is a requirement for the latter to be
achieved.  Correctly understanding and reconstructing every mail delivery
attempt is important so that the information in the database is accurate and
complete.  Improving the proportion of log lines correctly recognised is
the less difficult of the two, because it just requires new rules to be
written or existing rules to be extended.  Improving the proportion of
correctly understood and reconstructed mail delivery attempts is more
difficult and intrusive, because it requires adding or changing actions,
and it can be much harder to realise that a deficiency exists and needs to
be addressed.

@@

    \item Skip the log line if the ruleset does not contain any rules for
        the Postfix component that produced it.

    \item Try each rule for that Postfix component, then generic rules,
        until a recognising rule is found; if the log line is not
        recognised, issue a warning and move on to the next log line.

@@

Each Postfix component whose log lines are of interest must have at least
one rule, or all of its log lines will be silently skipped; in the extreme
case of an empty ruleset the parser would skip every log line.
\parsername{} skips log lines logged by programs that do not have any
associated rules, because there may be any number of log lines from other
programs in the log file, and some Postfix components do not produce any
log lines of interest.  \parsername{} does not parse log lines from
non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could easily be
extended to do so, if a method could be developed to correctly associate
such log lines with existing state table entries.  To correctly recognise
all log lines, there must be a rule to recognise each log line variant
produced by each Postfix component; if a log line is not recognised the
parser will issue a warning, to inform the user that they need to extend
their ruleset.  Each rule's regex should be as specific and precise as
possible, to ensure accurate parsing: a rule with a regex that matches zero
or more of any character will recognise every log line, but not in a
meaningful way; most log lines contain fixed strings, so this is not a
problem in practice.

Full coverage of log lines can be achieved without undue effort, yet once
achieved it requires maintenance.  Maintaining full coverage takes effort because
the set of log line variants changes over time, e.g.\ administrators add
restrictions with custom messages, \acronym{DNSBL} messages change, or
major releases of Postfix change log lines (usually by adding more
information).  Warnings are issued for any log lines that are not
recognised; no warnings are issued for unrecognised log lines while parsing
the \numberOFlogFILES{} log files, so it can be safely concluded that zero
false negatives arise.  False positives are harder to quantify, short of
examining each of the 60,721,709 log lines and checking that the correct
rule recognised it.  However, a random sample of 6039 log lines was parsed,
and the results manually verified by inspection to ensure that the correct
rule recognised each log line.  The sample was generated by running the
command: \texttt{perl -n -e \singlequote{}print if (rand 1 <
0.0001)\singlequote{} LOG\_FILES} to randomly extract roughly one log line
in every 10,000 (it actually extracted 0.00994\% instead of 0.01\%).  Each
log line was examined and the correct rule identified from the
\numberOFrules{} rules in the database; the correct rule was then compared
to the rule that recognised the log line when parsing.  The sample results
contained zero false positives, and this check has been automated to ensure
continued accuracy.  Based on these results, and how precise each rule's
regex is, the author is confident that zero false positives occur when
parsing the \numberOFlogFILES{} log files.

@@

The proportion of mail delivery attempts that are correctly understood and
reconstructed is much more difficult to determine accurately than the
proportion of log lines that are correctly recognised.  The parser can dump
its state tables in a human readable form; examining those tables with
reference to the log files is the best way to detect mails that were not
reconstructed properly; many of the complications discussed in
\sectionref{complications} were detected in this way.  \parsername{} issues
warnings when it detects any errors or discrepancies, e.g.\ when a queueid
is reused but the previous mail remains in the state tables, when a queueid
or \acronym{pid} is not found in the state tables, or when an entry in the
state tables does not include sufficient data to satisfy the database
schema.  The parser should produce few or no warnings during parsing, and
when finished parsing the state tables should only contain entries for
mails that have log lines in subsequent log files.  There will often be
warnings about a missing queueid or \acronym{pid} when parsing the first
few thousand log lines, because the earlier log lines for those connections
or mails are in a previous log file; loading saved state tables from the
previous log file will solve this problem.

@@

The state tables will contain entries for mails not yet delivered when the
parser finishes parsing the log file.  Ideally, those are the only entries
they will contain, though they may also contain mails whose initial log
lines are not contained in the log files.  Any other entries in the state
tables are evidence of either a failure in parsing, or an aberration in the
log files.  After parsing the \numberOFlogFILES{} log files, the state
tables contain 18 entries:

@@

    \item 1 connection that started only seconds before the log files ended
        and the mail had not yet been fully transferred from client to
        server.

@@

    \item \parsername{} performs many checks to detect known problems
        occurring, e.g.\ if a mail is missing required data.  No such
        warnings are produced during the test runs other than the five
        described above.

@@

    \item When dealing with the complications described in
        \sectionref{complications}, the solutions are as specific and
        restrictive as possible, with the goal of minimising the number of
        false positives.  In addition, the solution to the complication
        described in \sectionref{out of order log lines} imposes conditions
        that every reconstructed mail must comply with to be acceptable,
        not just the five mails that exhibit that complication.
